---
draft: false
keywords: ""
slug: project
title: "GDP Components over time"
author: "Architt"
output: html_document
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(tidyr)
library(janitor)
library(stringr)
```




>In this Project we Manipulated raw GDP data to create meaningful Charts

### Challenge 2: GDP components over time and among countries

At the risk of oversimplifying things, the main components of gross domestic product, GDP are personal consumption (C), business investment (I), government spending (G) and net exports (exports - imports). You can read more about GDP and the different approaches in calculating at the [Wikipedia GDP page](https://en.wikipedia.org/wiki/Gross_domestic_product).

The GDP data we will look at is from the [United Nations' National Accounts Main Aggregates Database](https://unstats.un.org/unsd/snaama/Downloads), which contains estimates of total GDP and its components for all countries from 1970 to today. We will look at how GDP and its components have changed over time, and compare different countries and how much each component contributes to that country's GDP. The file we will work with is [GDP and its breakdown at constant 2010 prices in US Dollars](http://unstats.un.org/unsd/amaapi/api/file/6) and it has already been saved in the Data directory. Have a look at the Excel file to see how it is structured and organised


```{r read_GDP_data, message = FALSE, error = FALSE}

UN_GDP_data  <-  read_excel(here::here("data", "Download-GDPconstant-USD-countries.xlsx"), 
                sheet="Download-GDPconstant-USD-countr", # Sheet name
                skip=2) # Number of rows to skip
glimpse(UN_GDP_data)
```

 The first thing you need to do is to tidy the data, as it is in wide format and you must make it into long, tidy format. Please express all figures in billions (divide values by `1e9`, or $10^9$), and you want to rename the indicators into something shorter.


```{r reshape_GDP_data,message = FALSE, error = FALSE}

#Change the code to longer
tidy_GDP_data  <-  UN_GDP_data %>% 
  pivot_longer(cols=-c(CountryID, Country, IndicatorName), names_to = 'year', values_to = 'GDP') %>% 
  
#Change the code to required format  
  mutate(GDP = GDP/1e9) %>% 
  pivot_wider(names_from = IndicatorName, values_from ="GDP") %>% 
  
#rename Indicator names to shorter names  
  rename('FinConExp' = 'Final consumption expenditure', 
         "Household_Expenditure" = "Household consumption expenditure (including Non-profit institutions serving households)", 
         "Government_Expenditure" = "General government final consumption expenditure", 
         "Gross_capital_formation" = "Gross capital formation", 
         "Gross_fixed_capital_formation" = 'Gross fixed capital formation (including Acquisitions less disposals of valuables)', 
         "Exports" = "Exports of goods and services",
         "Imports" = "Imports of goods and services",  
         "GDP" = "Gross Domestic Product (GDP)", 
         "AgricultureEtc" = "Agriculture, hunting, forestry, fishing (ISIC A-B)", 
         "MiningEtc" = "Mining, Manufacturing, Utilities (ISIC C-E)", 
         "Manufacturing" = 'Manufacturing (ISIC D)',        
         "Construction" = "Construction (ISIC F)", 
         "Wholesale, Retail, Restaurants" = 'Wholesale, retail trade, restaurants and hotels (ISIC G-H)', 
         "Transport, Communication" = 'Transport, storage and communication (ISIC I)', 
         "Other" = 'Other Activities (ISIC J-P)',
         "Total_Value" = "Total Value Added", 
         "InventoryChange" = 'Changes in inventories'

  )

```


We Will Sort the Data and then Reproduce the graph

```{r country_plot,message = FALSE, error = FALSE,fig.width= 10, fig.height= 6}
#Create a specific data table
country_list <- tidy_GDP_data %>% 
  filter(Country == 'India' | Country == "United States" | Country == "Germany" ) %>% 
  select(contains ("Country"), 
         contains("year"), 
         contains("Gross_capital_formation"), 
         contains("Exports"), 
         contains("Government_Expenditure"), 
         contains("Imports"), 
         contains("Household_Expenditure"), 
         contains("GDP"))

glimpse(country_list)
 
#plotting the table 
ggplot(country_list, aes(x = as.double(year))) +
  
#Adding the lines  
  geom_line(aes(y = Gross_capital_formation, 
            color = '#f88f88'), 
            size=1) +
  
  geom_line(aes(y = Exports, 
            color = '#b1b327'), 
            size=1) +
  geom_line(aes(y = Government_Expenditure,
            color = '#13c487'), 
            size=1) +
  geom_line(aes(y=Imports, 
            color = "#eb86f5"), 
            size = 1)+
  geom_line(aes(y=Household_Expenditure,
            color = "#17b6f6"), 
            size = 1) +
  
#Adjusting the legend
  scale_color_identity(name = "Components of GDP",
                       labels = c( "Goverment Expenditure",
                                   "Household Expenditure",
                                   "Exports",
                                   "Imports",
                                   "Gross Capital Formation"),
                       guide = "legend")+
  
  #Sorting by country
  facet_wrap(~Country) +
  
  #Adding theme
  theme_bw() +
  
  #Adjusting plot aspect ratio 
  theme(aspect.ratio=20/9) +
  
  #Making scale continous and removing expansion
  scale_x_continuous(limits = c(1970,2017), expand = c(0, 0)) +
  
  #Adding plot information
   labs(x = "", y="Billion USD$", 
       title = "GDP components over time", 
       subtitle="In constant 2010 USD",color = "Legend")

```


Secondly, recall that GDP is the sum of Household Expenditure (Consumption *C*), Gross Capital Formation (business investment *I*), Government Expenditure (G) and Net Exports (exports - imports). Even though there is an indicator `Gross Domestic Product (GDP)` in your dataframe, I would like you to calculate it given its components discussed above.

> What is the % difference between what you calculated as GDP and the GDP figure included in the dataframe?


```{r percentage_diff, message = FALSE, error = FALSE}

# Caculating the new GDP and then percentages 
new_country_list <- country_list %>% 
  mutate((GDP_CIGNE = Government_Expenditure + 
                      Household_Expenditure + Gross_capital_formation + Exports - Imports),
          GDP_diff = ((GDP_CIGNE-GDP)/GDP)*100,
          Gov_perc = (Government_Expenditure/GDP_CIGNE),
          Hos_perc = (Household_Expenditure/GDP_CIGNE),
          GCF_perc = (Gross_capital_formation/GDP_CIGNE),
          NE_perc = ((Exports - Imports)/GDP_CIGNE))


GDP_perc_table <- new_country_list %>% 
  select(contains ("Country"),
         contains("year"),
         contains ("GDP"),
         contains ("GDP_CIGNE"),
         contains("GDP_diff"))

glimpse(GDP_perc_table)

```

We sort the Data and Produce the plot

```{r plot_2, message = FALSE, error = FALSE,fig.width= 10, fig.height= 6}

#plotting the table 
ggplot(new_country_list, aes(x = as.double(year))) +
  
#Plotting the lines  
  
  geom_line(aes(y = Gov_perc, 
            col = '#f88f88'), 
            size=.8, 
            alpha = 0.7) +
  
  geom_line(aes(y = GCF_perc, 
            col = '#b1b327'),
            size=.8, 
            alpha = 0.7) +
  
  geom_line(aes(y=NE_perc, 
            col = "purple"), 
            size = .8, 
            alpha = 0.7)+
  
  geom_line(aes(y=Hos_perc, 
            col = "#17b6f6"),
            size = .8, 
            alpha = 0.7) +
  
#Adjusting the Legend  
  scale_color_identity(name = NULL,
                       labels = c("Household Expenditure",
                                  "Gross Capital Formation",
                                  "Goverment Expenditure",
                                  "Net Exports"),
                       guide = "legend")+
  
  #Sorting by country
  facet_wrap(~Country) +
  
  #adding theme
  theme_bw() +
  
  #Adjusting plot aspect ratio 
  theme(aspect.ratio=20/9) +
  
  #Making Scales continous, removing exapnsion, adding percentage
  scale_x_continuous(limits = c(1970,2017), 
                     expand = c(0, 0)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
  
  #adding Plot information
   labs(x = "", y="proportion", 
       title = "GDP and its breakdown at constant 2010 prices in US Dollars",
       caption = "Source: United Nations, https://unstats.un.org/unsd/snaama/Downloads") 

```

> What is this last chart telling you? Can you explain in a couple of paragraphs the different dynamic among these three countries? 

**The primary difference between the three countries is the state of economic development. Germany and the United States are developed economies, while India is a developing economy.**

**The level of Household Expenditure for Germany and the United States has been relatively stable as a proportion to the total GDP at approximately 60%. Meanwhile, India starkly contrasts with these two countries, as the proportion of its Household Expenditures has decreased from 82% to 55% over 50 years. The change started in 1990, when the Indian Economy was liberalised and became a more business-friendly nation, where individuals and companies saw more opportunities and diverted capital from Consumption to Investment (GCF).**

**In terms of GCF, Germany has experienced a decrease in proposition by almost 10% over the last 50 years. This can be attributed to the fact that post-WW2, Germany was heaving investing into the reconstruction of its economy. After the late 1990s, when East Germany was reunified with West Germany, Investment decreased to approximately 20%, about the same level as the United States, signalling that it became a healthy economy. The United States, on the other hand, has had relatively stable investment, excluding dips in investment due to business cycles, specifically economic boom in the early 2000s and the 2009 financial crisis.**

**Government expenditure differs by country, however, has remained relatively stable over the last 50 years. Every country has its own expectations of growth rate, and generally are able to manage their government budget as an expected percentage of GDP. Additionally, interest rates differed greatly across countries. In developing countries like India, the government's borrowing costs were high, particularly pre-1990s. Thus, the Indian government had to be prudent in their spending patterns. With the United States being the heart and soul of capitalism, over the years, more power has been transferred to markets. Thus, this may be the reason that government expenditure has been falling in the US.**

**Lastly, the level of net exports, defined as exports minus imports has varied across countries. In Germany, the proportion of net exports has increased significantly after the 2000s, starting from being a balanced importing and exporting nation to becoming an export-driven nation. It is possible that this change occurred due to increased goods mobility in the EU. Both India and the United States, on the other hand, have become predominantly net importers over the last few decades. This implies that the countries import a greater value of goods than they export. In India, this can be attributed to the market liberalisation. Meanwhile, the United States has the largest goods consumption in the world, which explains a growing influx in imports in recent years.**




