---
draft: false
keywords: ""
slug: work 
title: "About my Work"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  tidy = TRUE,
  size = "small"
)
```


```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(gapminder)  # gapminder dataset
library(here)
library(janitor)
```




### Academic Work
*Research*
During my Undergraduate, as my final project, my study group researched for [The Antecedents and Lead Indicators to a Global Financial Crisis.](https://drive.google.com/file/d/1t2JWmXFCOtEFaD7TxS-jLfmdS0NaOVWP/view?usp=sharing)

It was about critically analyze the various investment avenues and their
interdependencies with the stock market, in order to establish certain movements in them as
indicators for the stock market crash.

Multiple variables had been analyzed, with the dependent variable being the S&P500 index,
as an assumption that the US has been the epicenter for most crises has been made.
A descriptive research with secondary data was carried out, correlational tools were used on
data for the last 40 years, covering 3 major crises: the 1980s recession, the dot-com bubble,
and the sub-prime mortgage crisis.
The research was successful at establishing relationships between the variables.

*Project*

we have also studied[Germany's economic Influence over Europe post 1970's.](https://drive.google.com/file/d/1vOH4cOUfC-CMQCax8urPoblYdzCZdCc9/view?usp=sharing)

In this project, we analysed the German influence in the peace era post world War 2 and its role in the Economic and Monetary union of Europe 

